Combining Object-Oriented Paradigm and Controlled Natural Language for Requirements Specification

### 2021-08-01 article.tex
acmart

\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{silence}
  \WarningFilter*{latex}{Font shape `OT1/zi4/m/it' undefined}
\usepackage[table]{xcolor} % to highlight rows in tables
\usepackage[underline=true,rounded corners=false]{pgf-umlsd} % for UML sequence diagrams
\usepackage{inconsolata} % for nicer TT font family
\usepackage[square]{syntax} % for BNF grammar
  \setlength{\grammarparsep}{1pt}
  \setlength{\grammarindent}{2em}
\usepackage{minted} % code snippets
  \newminted{text}{linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small}
\usepackage{array} % complex structure of TABULAR-s
\usepackage{paralist} % for inline itemized lists env{inparaenum}
\usepackage{natbib} % for customization of bibliography headers
\usepackage{adjustbox}
  \newcommand{\code}[1]{\adjustbox{padding=3pt 1pt 3pt 1pt, cframe=gray!50, bgcolor=gray!10}{\texttt{#1}}}
\usepackage{tikz}
  \usetikzlibrary{shapes,arrows,shadows}
  \usetikzlibrary{decorations.pathmorphing}
  \usetikzlibrary{positioning,fit}
  \usetikzlibrary{calc}
\usepackage[hyperfootnotes=true,unicode]{hyperref} % hyper links
  \hypersetup{colorlinks=false}

@@

\begin{document}
\tolerance=1500
\raggedbottom

\title{Combining Object-Oriented Paradigm and Controlled Natural Language for Requirements Specification}
\author{Yegor Bugayenko\\yegor256@gmail.com}
\maketitle

\begin{abstract} Natural language is the dominant form of writing software requirements.
Its essential ambiguity causes inconsistency of requirements,
which leads to scope creep. On the other hand, formal requirements
specification notations such as Z, \nospell{Petri Nets}, SysML, and others are
difficult to understand by non-technical project stakeholders. They often
become a barrier between developers and requirements providers. The article
presents a controlled natural language that looks like English but
is a strongly typed object-oriented language compiled to
UML/XMI. Thus, it is easily understood, at the same time, by non-technical
people, programmers, and computers. Moreover, it is formally verifiable and
testable. It was designed, developed, and tested in three commercial
software projects in order to validate the assumption that object-oriented
design can be applied to requirements engineering at the level of
specifications writing. The article outlines key features of the language and summarizes
the experience obtained during its practical application. \end{abstract}

@@

Software Requirements Specification (SRS) should be correct, unambiguous,
complete, consistent, ranked for importance and/or stability, verifiable,
modifiable, and traceable~\cite{ieee830}. Many times, it was proven that an SRS
document that lacks some of these properties has a high likelihood of
becoming a root cause of scope and schedule problems in a
project~\cite{wiegers03,ears09,chen09}.

Natural Language (NL) is the main presentation means in industrial
requirements documents~\cite{kof10,sinha10}, mostly due to its flexibility.
Very often, contributors to requirements specifications are not experts in
requirements engineering (RE). Impreciseness and flexibility of NL, being
its advantage in informal human communications, turns into a serious drawback
in RE. Problems that appear in NL requirement documents include
ambiguity, vagueness, complexity, duplication, wordiness, and
\nospell{untestability}~\cite{ears09}.

To overcome problems associated with NL, some experts advocate the use of
other notations for the specification of requirements, such as
Z~\cite{woodcock96}, SysML, UML~\cite{uml}, Petri Nets, and others. However,
use of any of these non-textual notations often requires complex translation
of the source requirements, which can introduce further errors. Such
translation of requirements can serve to create an extra ``language
barrier'' between developers and stakeholders. There is also a training
overhead associated with the introduction of many notations~\cite{ears09}.

A desired solution to the outlined problem would be a Controlled Natural
Language (CNL), which would look like NL, but be precise,
unambiguous, and consistent. A few CNLs were introduced and developed over
the last decades. Closest to our work are Attempt to Controlled English
(ACE)~\cite{ace06}, Simplified Technical English (STE)~\cite{ste05}, Easy
Approach to Requirements Syntax (EARS)~\cite{ears09}, Boeing's Computer
Processable Language~\cite{clark05}, and \nospell{Schwitter's Processable
ENGlish}~\cite{peng}. However, there are a few critical drawbacks that exist,
to some degree, in all of these CNLs.

@@

Third, most do not allow uncertainty, while requirements development
and requirements management processes need to be able to mark
requirements as uncertain (also known as ``TBD''). Requirements engineers
need to have an ability to add a requirement into scope at one iteration
and refine it later in a few iterations. This process, called ``progressive
elaboration'' of requirements~\cite{pmbok4}, must be supported by a CNL.

@@

It is an object-oriented CNL that resembles English and, because of that,
can be read and understood by compilers, system analysts, end-users, and
business people. Fig.~\ref{fig:example} presents an example of a short
requirements definition document that describes a calculator that asks its
user for two floating point numbers and returns the result of their
division.

\begin{figure}
\begin{textcode}
SuD includes: user as User.
Fraction is a "math calculator".
Fraction needs:
numerator as Float, and
denominator as Float.
UC1 where SuD divides two numbers:
1. The user creates Fraction (a fraction);
2. The fraction "calculates" Float (a quotient);
3. The user "receives results" using the quotient.
UC1/2 when "division by zero":
1. The user "fails" using "can't divide by zero".
\end{textcode}
\caption{Software Requirements Specification of an example math calculator}
\label{fig:example}
\end{figure}

The example has some uncertainty expressed with plain English text sentences
in double quotes. This is an important feature that distinguishes the designed language from
other CNLs. Every class and method may be specified either formally or
by so-called ``informal'' text in double quotes. In NL requirements
specifications, such uncertainties are usually marked with TBD (``to be
determined'') placeholders~\cite{moynihan00}. Java equivalent of them are
\code{TODO} comments. Informal texts in double quotes are also used to
make references from functional requirements to customer needs, interface
specifications, supplementary documents, etc.

\code{SuD} (System under Development) is a top level class that encapsulates
all other objects and may not be ``created'' (similar to a singleton in OOP).
\code{Float} is a built-in class, as well as \code{Text}, \code{Integer}, and
a few others. Classes have methods, which are defined as use cases with a
single main flow and a few alternative flows~\cite{cockburn01}.

In the example, use case \code{UC1} is a method of class \code{SuD} and its
name is ``divides two numbers.'' Every flow step in a use case can either
\begin{inparaenum}[\itshape a\upshape)] \item call other method, or \item
create (instantiate) an object. \end{inparaenum} In the example, step
\code{1} of \code{UC1} is creating a new object \code{fraction} of class
\code{Fraction}, while step \code{2} is calling a method ``calculates'' of
\code{fraction} in order to get back an object named \code{quotient} of
class \code{Float}.

@@

Fig.~\ref{fig:java} contains a close equivalent program in Java.

\begin{figure}
\inputminted[linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small]{java}{sample.java}
\caption{Java equivalent of Requs requirements document from Fig.~\ref{fig:example}}
\label{fig:java}
\end{figure}

@@

In Fig.~\ref{fig:example}, \code{SuD}, \code{Fraction}, \code{Float}, and
\code{User} are classes. They start with capital letters and contain only
letters (both lowercase and capital). \code{user}, \code{fraction}, and
\code{quotient} are variables bound to instances/objects of classes.

Besides, there are a few reserved words, which may have special meaning in
certain places of the text, including \code{the}, \code{a}, \code{includes},
\code{requires}, etc. (A full list of them is defined below in Section~\ref{sec:syntax}.)

@@

Class declaration is a clause that declares a class as a sub-class of
another class or as a standalone class. The central part of the clause is
an \code{is a} term:

@@

The second line in Fig.~\ref{fig:example} declares class \code{Fraction}
without any super-class but with an informal description \code{"math
calculator"}.

@@

Fig.~\ref{fig:example} defines constructor arguments only for
\code{Fraction} class. Class \code{User} don't have any arguments. Class
\code{SuD} is a pre-defined class without arguments. Class \code{Float} is a
pre-defined class with one argument.

Class without arguments in constructor normally means that it represents an
entity outside of system scope, like \code{User}.

Cardinality of associations between objects is configured by the \code{-s}
prefix at the end of variable name. There are only
two possible relationships types supported: one-to-one and one-to-many.

@@

For traceability, all methods have unique names across the entire
SRS starting with \code{UC} (use case). In other words,
methods are use cases owned by classes.

@@

Every method call in every use case should either
\begin{inparaenum}[\itshape a\upshape)]
\item be informal (the 8th, 9th, and 11th lines in the example at Fig.~\ref{fig:example}),
\item formally refer to the method declared in a class, or
\item construct an instance of a class using a built-in method \code{creates} (the 7th line in the example).
\end{inparaenum}

@@

Every time an object is instantiated using the \code{creates} method,
the compiler checks that all required parameters of required classes are
supplied. Again, it is a very important validation that prevents inconsistency
during requirements maintenance and refinement. In the example, class
\code{Fraction} requires two parameters to be instantiated. Any attempt to
create a fraction with just one parameter or two parameters of classes other
than \code{Float} would be rejected by the compiler.

@@

A variable is declared and bound to its value by means of an article
\code{a}. The seventh line of the example declares a variable
\code{fraction}, which is later used with article \code{the} (the eighth
line). If, by mistake, variable declaration is removed from the first step
of the use case or the entire step is deleted, compilation will fail.
Consistency of the document is ensured by this constraint.

@@

Similar to the check of parameter classes, every returned object is verified
for compliance with the class expected. The eighth line of the example
expects method ``calculates'' of class \code{Fraction} to return an object
of class \code{Float}. The method is still informal and is not defined in
the requirements document, but when a requirements expert decides to make it
formal and specify its details in a new use case, she has to make sure it
returns an instance of class \code{Float}. Otherwise, the compiler would
complain with an error.

@@

Every method in the object-oriented way may throw an exception, using a
build-in method \code{Fail since}. The method has one parameter of class
\code{Text}, which explains the reason. In the example in
Fig.~\ref{fig:example}, method ``calculates'' of class \code{Fraction} may
throw an exception and it is handled by an alternative flow. In the language, all
exceptions are checked, which means that they should be handled explicitly,
by means of alternative flows of use cases.

@@

When semantic analysis is done the specification looks like an object of
class \code{SuD} that encapsulates other objects and methods. This object
hierarchy is convertible to UML diagrams~\cite{uml}. Specification in
Fig.~\ref{fig:example} would be converted to the following UML diagrams
(to save space, the list is limited to the three most interesting
diagrams):

\begin{itemize}
\item Class diagram, in Fig.~\ref{fig:uml-class}
\item Object diagram, in Fig.~\ref{fig:uml-object}
\item Sequence diagram of ``divides two numbers'' method, in Fig.~\ref{fig:uml-seq}
\end{itemize}

@@

Every UML diagram is a formal interoperable document in XMI~\cite{xmi} that
can be rendered in web, in \LaTeX, in PDF, or translated to Java or any
other object-oriented language. Fig.~\ref{fig:xmi} shows an example XMI
for the class diagram.

\begin{figure}
\inputminted[linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small]{xml}{sample.xmi}
\caption{XMI representation of a UML diagram; a sample part}
\label{fig:xmi}
\end{figure}

@@

In Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
methods in total and three of them are defined informally.

A method is formally defined when its signature consists of one specially
reserved verb \code{creates} and an optional list of constructor arguments,
as in the seventh line of Fig.~\ref{fig:example}. A formally defined
method doesn't have any ambiguity since it is absolutely clear what is
expected as its input and output. All it does is instantiate a new object
and bind it to a variable. In Java that would mean calling a constructor.

@@

\begin{table}
\begin{tabular}{>{\raggedright}p{14em}rrr}
                                            & $p_1$ & $p_2$ & $p_3$ \\
\hline
Total classes                                 & 39 & 12 & 19 \\
Total methods                                 & 19 & 7 & 24 \\
Total Java classes                                  & 80 & 75 & 295 \\
\rowcolor{lightgray!50}
Ambiguity achieved                                  & 0.80 & 0.95 & 0.68 \\
Work spent on RE, staff-months                      & 0.53 & 0.21 & 0.86 \\
Total SRS contributors                              & 4 & 7 & 2 \\
Non-empty lines of Java code                        & 9.5K & 10K & 32K \\
Project duration, months                            & 7 & 26 & 11 \\
Project budget, staff-months                        & 5.5 & 7.5 & 9.5 \\
\hline
\end{tabular}
\caption{Empirically collected data from three commercial projects}
\label{table:numbers}
\end{table}

@@

Third, developing a requirements specification with the language requires RE
engineers to understand key principles of ``object thinking''~\cite{west04},
like ``everything is an object'' and ``objects expose behavior, not state.''
Initial training was required in two projects.

>>>>

\documentclass[sigplan,10pt,nonacm=true]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[underline=true,rounded corners=false]{pgf-umlsd} % for UML sequence diagrams
\usepackage[square]{syntax} % for BNF grammar
  \setlength{\grammarparsep}{1pt}
  \setlength{\grammarindent}{2em}
\usepackage{array} % complex structure of TABULAR-s
\usepackage{paralist} % for inline itemized lists env{inparaenum}
\usepackage{natbib} % for customization of bibliography headers
\usepackage{ffcode}
\usepackage{tikz}
  \usetikzlibrary{shapes,arrows,shadows}
  \usetikzlibrary{decorations.pathmorphing}
  \usetikzlibrary{positioning,fit}
  \usetikzlibrary{calc}
\usepackage{href-ul} % hyper links

@@

\begin{document}

\title{Combining Object-Oriented Paradigm and Controlled Natural Language for Requirements Specification}
\author{Yegor Bugayenko\\yegor256@gmail.com}
\begin{abstract} Natural language is the dominant form of writing software requirements.
Its essential ambiguity causes inconsistency of requirements,
which leads to scope creep. On the other hand, formal requirements
specification notations such as Z, \nospell{Petri Nets}, SysML, and others are
difficult to understand by non-technical project stakeholders. They often
become a barrier between developers and requirements providers. The article
presents a controlled natural language that looks like English but
is a strongly typed object-oriented language compiled to
UML/XMI. Thus, it is easily understood, at the same time, by non-technical
people, programmers, and computers. Moreover, it is formally verifiable and
testable. It was designed, developed, and tested in three commercial
software projects in order to validate the assumption that object-oriented
design can be applied to requirements engineering at the level of
specifications writing. The article outlines key features of the language and summarizes
the experience obtained during its practical application. \end{abstract}
\maketitle

@@

Software Requirements Specification (SRS) should be correct, unambiguous,
complete, consistent, ranked for importance and/or stability, verifiable,
modifiable, and traceable~\citep{ieee830}. Many times, it was proven that an SRS
document that lacks some of these properties has a high likelihood of
becoming a root cause of scope and schedule problems in a
project~\citep{wiegers03,ears09,chen09}.

Natural Language (NL) is the main presentation means in industrial
requirements documents~\citep{kof10,sinha10}, mostly due to its flexibility.
Very often, contributors to requirements specifications are not experts in
requirements engineering (RE). Impreciseness and flexibility of NL, being
its advantage in informal human communications, turns into a serious drawback
in RE. Problems that appear in NL requirement documents include
ambiguity, vagueness, complexity, duplication, wordiness, and
\nospell{untestability}~\citep{ears09}.

To overcome problems associated with NL, some experts advocate the use of
other notations for the specification of requirements, such as
Z~\citep{woodcock96}, SysML, UML~\citep{uml}, Petri Nets, and others. However,
use of any of these non-textual notations often requires complex translation
of the source requirements, which can introduce further errors. Such
translation of requirements can serve to create an extra ``language
barrier'' between developers and stakeholders. There is also a training
overhead associated with the introduction of many notations~\citep{ears09}.

A desired solution to the outlined problem would be a Controlled Natural
Language (CNL), which would look like NL, but be precise,
unambiguous, and consistent. A few CNLs were introduced and developed over
the last decades. Closest to our work are Attempt to Controlled English
(ACE)~\citep{ace06}, Simplified Technical English (STE)~\citep{ste05}, Easy
Approach to Requirements Syntax (EARS)~\citep{ears09}, Boeing's Computer
Processable Language~\citep{clark05}, and \nospell{Schwitter's Processable
ENGlish}~\citep{peng}. However, there are a few critical drawbacks that exist,
to some degree, in all of these CNLs.

@@

Third, most do not allow uncertainty, while requirements development
and requirements management processes need to be able to mark
requirements as uncertain (also known as ``TBD''). Requirements engineers
need to have an ability to add a requirement into scope at one iteration
and refine it later in a few iterations. This process, called ``progressive
elaboration'' of requirements~\citep{pmbok4}, must be supported by a CNL.

@@

It is an object-oriented CNL that resembles English and, because of that,
can be read and understood by compilers, system analysts, end-users, and
business people. The Fig.~\ref{fig:example} presents an example of a short
requirements definition document that describes a calculator that asks its
user for two floating point numbers and returns the result of their
division.

\begin{figure}
\begin{ffcode}
SuD includes: user as User.
Fraction is a "math calculator".
Fraction needs:
numerator as Float, and
denominator as Float.
UC1 where SuD divides two numbers:
1. The user creates Fraction (a fraction);
2. The fraction "calculates" Float
  (a quotient);
3. The user "receives results" using
  the quotient.
UC1/2 when "division by zero":
1. The user "fails" using "can't
  divide by zero".
\end{ffcode}
\caption{Software Requirements Specification of an example math calculator}
\label{fig:example}
\end{figure}

The example has some uncertainty expressed with plain English text sentences
in double quotes. This is an important feature that distinguishes the designed language from
other CNLs. Every class and method may be specified either formally or
by so-called ``informal'' text in double quotes. In NL requirements
specifications, such uncertainties are usually marked with TBD (``to be
determined'') placeholders~\citep{moynihan00}. Java equivalent of them are
\ff{TODO} comments. Informal texts in double quotes are also used to
make references from functional requirements to customer needs, interface
specifications, supplementary documents, etc.

\ff{SuD} (System under Development) is a top level class that encapsulates
all other objects and may not be ``created'' (similar to a singleton in OOP).
\ff{Float} is a built-in class, as well as \ff{Text}, \ff{Integer}, and
a few others. Classes have methods, which are defined as use cases with a
single main flow and a few alternative flows~\citep{cockburn01}.

In the example, use case \ff{UC1} is a method of class \ff{SuD} and its
name is ``divides two numbers.'' Every flow step in a use case can either
\begin{inparaenum}[\itshape a\upshape)] \item call other method, or \item
create (instantiate) an object. \end{inparaenum} In the example, step
\ff{1} of \ff{UC1} is creating a new object \ff{fraction} of class
\ff{Fraction}, while step \ff{2} is calling a method ``calculates'' of
\ff{fraction} in order to get back an object named \ff{quotient} of
class \ff{Float}.

@@

The Fig.~\ref{fig:java} contains a close equivalent program in Java.

\begin{figure}
\begin{ffcode}
class User {
  Fraction createsFraction() {
  }
  void receivesResult(Float quotient) {
  }
  void fails(String message) {
  }
}
class Fraction {
  private final float numerator;
  private final float denumerator;
  Fraction(float num, float denum) {
    this.numerator = num;
    this.denumerator = denum;
  }
  Float quotient() throws DivisionByZero {
  }
}
class SuD {
  private final User user;
  void dividesTwoNumbers() {
    Fraction fraction = this.user.createsFraction();
    try {
      this.user.receivesResult(
        fraction.quotient()
      );
    } catch (DivisionByZero ex) {
      this.user.fails(
        "denominator can't be zero"
      );
    }
  }
}
\end{ffcode}
\caption{Java equivalent of Requs requirements document from the Fig.~\ref{fig:example}}
\label{fig:java}
\end{figure}

@@

In the Fig.~\ref{fig:example}, \ff{SuD}, \ff{Fraction}, \ff{Float}, and
\ff{User} are classes. They start with capital letters and contain only
letters (both lowercase and capital). \ff{user}, \ff{fraction}, and
\ff{quotient} are variables bound to instances/objects of classes.

Besides, there are a few reserved words, which may have special meaning in
certain places of the text, including \ff{the}, \ff{a}, \ff{includes},
\ff{requires}, etc. (A full list of them is defined below in Section~\ref{sec:syntax}.)

@@

Class declaration is a clause that declares a class as a sub-class of
another class or as a standalone class. The central part of the clause is
an \ff{is a} term:

@@

The second line in the the Fig.~\ref{fig:example} declares class \ff{Fraction}
without any super-class but with an informal description
\ff{"math calculator"}.

@@

The Fig.~\ref{fig:example} defines constructor arguments only for
\ff{Fraction} class. Class \ff{User} don't have any arguments. Class
\ff{SuD} is a pre-defined class without arguments. Class \ff{Float} is a
pre-defined class with one argument.

Class without arguments in constructor normally means that it represents an
entity outside of system scope, like \ff{User}.

Cardinality of associations between objects is configured by the \ff{-s}
prefix at the end of variable name. There are only
two possible relationships types supported: one-to-one and one-to-many.

@@

For traceability, all methods have unique names across the entire
SRS starting with \ff{UC} (use case). In other words,
methods are use cases owned by classes.

@@

Every method call in every use case should either
\begin{inparaenum}[\itshape a\upshape)]
\item be informal (the 8th, 9th, and 11th lines in the example at the Fig.~\ref{fig:example}),
\item formally refer to the method declared in a class, or
\item construct an instance of a class using a built-in method \ff{creates} (the 7th line in the example).
\end{inparaenum}

@@

Every time an object is instantiated using the \ff{creates} method,
the compiler checks that all required parameters of required classes are
supplied. Again, it is a very important validation that prevents inconsistency
during requirements maintenance and refinement. In the example, class
\ff{Fraction} requires two parameters to be instantiated. Any attempt to
create a fraction with just one parameter or two parameters of classes other
than \ff{Float} would be rejected by the compiler.

@@

A variable is declared and bound to its value by means of an article
\ff{a}. The seventh line of the example declares a variable
\ff{fraction}, which is later used with article \ff{the} (the eighth
line). If, by mistake, variable declaration is removed from the first step
of the use case or the entire step is deleted, compilation will fail.
Consistency of the document is ensured by this constraint.

@@

Similar to the check of parameter classes, every returned object is verified
for compliance with the class expected. The eighth line of the example
expects method ``calculates'' of class \ff{Fraction} to return an object
of class \ff{Float}. The method is still informal and is not defined in
the requirements document, but when a requirements expert decides to make it
formal and specify its details in a new use case, she has to make sure it
returns an instance of class \ff{Float}. Otherwise, the compiler would
complain with an error.

@@

Every method in the object-oriented way may throw an exception, using a
build-in method \ff{Fail since}. The method has one parameter of class
\ff{Text}, which explains the reason. In the example in
the Fig.~\ref{fig:example}, method ``calculates'' of class \ff{Fraction} may
throw an exception and it is handled by an alternative flow. In the language, all
exceptions are checked, which means that they should be handled explicitly,
by means of alternative flows of use cases.

@@

When semantic analysis is done the specification looks like an object of
class \ff{SuD} that encapsulates other objects and methods. This object
hierarchy is convertible to UML diagrams~\citep{uml}. Specification in
the Fig.~\ref{fig:example} would be converted to the following UML diagrams
(to save space, the list is limited to the three most interesting
diagrams):

\begin{itemize}
\item Class diagram, in the Fig.~\ref{fig:uml-class}
\item Object diagram, in the Fig.~\ref{fig:uml-object}
\item Sequence diagram of ``divides two numbers'' method, in the Fig.~\ref{fig:uml-seq}
\end{itemize}

@@

Every UML diagram is a formal interoperable document in XMI~\citep{xmi} that
can be rendered in web, in \LaTeX, in PDF, or translated to Java or any
other object-oriented language. The Fig.~\ref{fig:xmi} shows an example XMI
for the class diagram.

\begin{figure}
\begin{ffcode}
<uml:Model xmi:version="2.1"
  xmlns:uml
    ="http://www.eclipse.org/uml2/2.0.0/UML"
  xmlns:xmi
    ="http://schema.omg.org/spec/XMI/2.1">
  <packagedElement name="SuD"
    xmi:type="uml:Class">
    <ownedOperation name="divides two numbers"
      xmi:type="uml:Operation">
      <ownedParameter direction="return"
        xmi:type="uml:Parameter"/>
    </ownedOperation>
  </packagedElement>
  <packagedElement name="User"
    xmi:type="uml:Class">
    <ownedOperation name="creates fraction"
      xmi:type="uml:Operation">
      <ownedParameter direction="return"
        xmi:type="uml:Parameter"/>
    </ownedOperation>
  </packagedElement>
  [...]
</uml:Model>
\end{ffcode}
\caption{XMI representation of a UML diagram; a sample part}
\label{fig:xmi}
\end{figure}

@@

In the Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
methods in total and three of them are defined informally.

A method is formally defined when its signature consists of one specially
reserved verb \ff{creates} and an optional list of constructor arguments,
as in the seventh line of the Fig.~\ref{fig:example}. A formally defined
method doesn't have any ambiguity since it is absolutely clear what is
expected as its input and output. All it does is instantiate a new object
and bind it to a variable. In Java that would mean calling a constructor.

@@

\begin{table}
\begin{tabular}{>{\raggedright}p{14em}rrr}
                                            & $p_1$ & $p_2$ & $p_3$ \\
\hline
Total classes                                 & 39 & 12 & 19 \\
Total methods                                 & 19 & 7 & 24 \\
Total Java classes                                  & 80 & 75 & 295 \\
Ambiguity achieved                                  & 0.80 & 0.95 & 0.68 \\
Work spent on RE, staff-months                      & 0.53 & 0.21 & 0.86 \\
Total SRS contributors                              & 4 & 7 & 2 \\
Non-empty lines of Java code                        & 9.5K & 10K & 32K \\
Project duration, months                            & 7 & 26 & 11 \\
Project budget, staff-months                        & 5.5 & 7.5 & 9.5 \\
\hline
\end{tabular}
\caption{Empirically collected data from three commercial projects}
\label{table:numbers}
\end{table}

@@

Third, developing a requirements specification with the language requires RE
engineers to understand key principles of ``object thinking''~\citep{west04},
like ``everything is an object'' and ``objects expose behavior, not state.''
Initial training was required in two projects.
